Co-moving stars in \textsl{Gaia DR1}

### 2016-10-19 paper/ms.tex
update results and figures

The primary data set used in this \documentname\ is the Tycho-Gaia Astrometric
Solution (\tgas), released as a part of Data Release 1 (\dr) of the Gaia mission
(\citealt{many}). \tgas\ contains astrometric measurements (sky position,
parallax, and proper motions) and associated covariance matrices for a large
fraction of the \project{Tycho-2} catalog (\citealt{tycho2}) with median
astrometric precision comparable to that of the \project{Hipparcos} catalog
($\approx 0.3~{\rm mas}$; \citealt{hipparcos}). In terms of parallax
signal-to-noise ($\snr = \varpi/\sigma_\varpi$), the \tgas\ catalog contains
42385 high-precision stars with $\snr > 32$: here we visualize quantities
computed from this high-precsision sample to emphasize the quality of the data.
In the sections that follow, we construct a statistical model that propagates
the non-trivial uncertainties in the data to our beliefs about the likelihood
that a given pair are co-moving.

@@

\section{Results 1: A catalog of candidate wide binaries}

>>>>

The primary data set used in this \documentname\ is the Tycho-Gaia Astrometric
Solution (\tgas), released as a part of Data Release 1 (\dr) of the Gaia mission
\citep{2016arXiv160904172G,2016arXiv160904303L}.
\tgas\ contains astrometric measurements (sky position,
parallax, and proper motions) and associated covariance matrices for a large
fraction of the \project{Tycho-2} catalog \citep{2000A&A...355L..27H} with median
astrometric precision comparable to that of the \project{Hipparcos} catalog
\citep[$\approx 0.3~{\rm mas}$;][]{2007ASSL..350.....V}. In terms of parallax
signal-to-noise ($\snr = \varpi/\sigma_\varpi$), the \tgas\ catalog contains
42385 high-precision stars with $\snr > 32$: here we visualize quantities
computed from this high-precision sample to emphasize the quality of the data.
In the sections that follow, we construct a statistical model that propagates
the non-trivial uncertainties in the data to our beliefs about the likelihood
that a given pair of stars is co-moving.

@@

\section{Results 1: A catalog of candidate wide binaries}
In this section, we examine the distribution of likelihood ratios
$\ln \mathcal{L}_1 /\mathcal{L}_2$, and come up with a reasonable cut
for this quantities to select co-moving pairs from candidate pairs list.
We present the statistics of the resulting co-moving pairs.
We also match the pairs with known open clusters in the Milky Way to
provide information on cluster membership.

Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all 2.6 million
candidate pairs.
Because we do not carry over the normalizations of all probabilities involved
in $L_1$ and $L_2$, the absolute values of the likelihood ratio is not
a meaningful quantity.
Given our crude way of coming up with candidate pair list, we expect
a huge number of pairs that are not genuinely co-moving, compared to
the smaller number of pairs that are.
In order to gauge an appropriate cut on the likelihood ratios,
we compute the likelihood ratios for 200,000 random pairs of stars with the same
signal-to-noise ratio cut as candidate pairs.
Additionally, we select from candidate pair list, those with very small
difference in tangential velocities ($\Delta v_t < 0.5$~km/s) in
two minimum signal-to-noise ratio\footnote{lower of the two stars in a pair}
bins (S/N$>32$ and S/N$<32$) that are highly probable
to be genuine co-moving pairs.
As expected, we see that the likelihood ratio $\ln \mathcal{L}_1 /\mathcal{L}_2$
distribution peaks at higher values, favoring hypothesis 1, for highly probable
co-moving pairs than the random pairs.
We also see that the values are even higher for higher minimum signal-to-noise ratio pairs.

Based on the comparison with random pairs, we can come up with a cut
on the likelihood ratios that reduces the cumulative contamination due to random pairs
to a given threshold.
The value of the cut, the resulting number of pairs, and
its fraction of the total number of candidate pairs
depending on the contamination rate are listed in
Table~\ref{tab:candidatecounts}.
Because the likelihood ratio $\ln \mathcal{L}_1 /\mathcal{L}_2$ asymtotically
is chi-square distribution, the power-law fall off makes the number of resulting
pairs sensitive to the cut.

Once we have decided the co-moving pairs from candidate pair list,
we can count how many co-moving neighbors a star has.
This is presented in Figure~\ref{fig:hist_Nneighbors}.
It is clear from this figure that we have many aggregates of co-moving stars, most
likely an open cluster or a moving group. We also see that there is still
a large number of stars which only one or two co-moving neighbors.
In order to select pairs that are not associated with any other pair,
we separate a subset of co-moving pairs whose member is a mutually exclusive
co-moving neighbor to the other member (``mutually exclusive pairs'').
The number of mutually exclusive pairs is listed on the last column of
Table~\ref{tab:candidatecounts}.

\begin{table}[tbh]
  \caption{Number of pairs}
  \label{tab:candidatecounts}
  \centering

  \begin{tabular}{l|c|c|c|r}
  \hline

  \hline
  \textbf{Contamination} & \textbf{$\ln \mathcal{L}_1 /\mathcal{L}_2$ cut} &
    \textbf{Count} & \textbf{Fraction} & \textbf{$N_\textrm{me}$} \\
  \hline
    0.01    \% & 5.33 & 92458 & 0.462 \% & 5055 \\
    0.001   \% & 6.53 & 13354 & 0.067 \% & 1549 \\
    0.0001  \% & 7.29 &  2832 & 0.014 \% & 582  \\
    0.00001 \% & 7.35 &  2495 & 0.012 \% & 532  \\
  \hline
  \end{tabular}
\end{table}

\begin{figure*}[p]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/likelihoodratios.pdf}
  \end{center}
  \caption{%
    Density histogram of likelihood ratios for candidate pairs and random pairs.
    We also show the distribution for highly probable genuine co-moving pairs with
    $\Delta v_t < 0.5$~km/s in high ($S/N>32$) and low ($S/N<32$) signal-to-noise ratio
    bins. In both $S/N$ bins, the highly probable co-moving pairs are clearly
    separated from random pairs while the distribution of $\ln \mathcal{L}_1 /\mathcal{L}_2$
    for candidate pairs is expected to be a combination of
    a huge number of ``random'' pairs that are not actually co-moving, and
    a smaller number of genuine co-moving pairs with varying $S/N$.
    \label{fig:likelihoodratios}}
\end{figure*}

We now examine the separation distribution of co-moving pairs in
Figure~\ref{fig:hist_separation}.

Finally, we show the distribution co-moving pairs in galactic longitude and
distance in Figure~\ref{fig:glon_d_pairlines}.
In the top panel, we show all co-moving pairs with contamination rate of $0.01~\%$
(undersampled by a factor of 4 in order to unclutter the plot) as connected line segments.
We find many aggregates of co-moving pairs, which coincides with the positions of
known Milky Way star clusters \citep{Kharchenko:2016aa} indicated by red circles.
Clumps of co-moving pairs at $l=300-360$~degrees and $d=100-200$~pc correspond
to the location of OB associations
Upper Scorpius, Upper Centaurus Lupus, and Lower Centaurus Crux
\citep{de-Zeeuw:1999aa}.
The strong correlation with known co-moving structures disappears when we only
consider mutually exclusive pairs (bottom panel).

\begin{figure*}[p]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/hist_Nneighbors.pdf}
  \end{center}
  \caption{%
    Histogram of the number of co-moving neighbors to a star for varying
    degrees of contamination due to random pairs.
    \todo{SMOH: What is the optimal bins for this histogram to separate
    small number cases e.g., binary case (N=1) appropriately?}
    \label{fig:hist_Nneighbors}}
\end{figure*}

\begin{figure*}[p]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/hist_separation.pdf}
    \includegraphics[width=\textwidth]{figures/hist_separation_mutexc.pdf}
  \end{center}
  \caption{%
    Top: Separation distribution of co-moving pairs of stars for varying degrees
    of contamination due to random pairs. The distribution peaks around 10~pc,
    which is similar to the typical tidal radius of clusters in the Milky Way
    \citep{Kharchenko:2013aa}.
    Bottom: Same as the previous plot but only for mutually exclusive pairs.
    \label{fig:hist_separation}}
\end{figure*}

\begin{figure*}[p]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/glon_d_pairlines.pdf}
    \includegraphics[width=\textwidth]{figures/glon_d_pairlines_mutexc.pdf}
  \end{center}
  \caption{%
    Top: Distribution of co-moving pairs in galactic longitude and
    distance. Each co-moving pair is connected by a line.
    Red circles mark the locations of 135 known Milky Way open clusters within
    $d<600$~pc from \citealt{Kharchenko:2016aa} while
    stars in known OB associations \citep{de-Zeeuw:1999aa} are colored
    by which association they belong.
    We see that some of the aggregates of co-moving pairs are clearly
    associated with known open clusters, and OB associations.
    Bottom: Same as the previous plot but only for mutually exclusive pairs.
    The mutually exclusive pairs no longer traces any known co-moving structures
    strongly.
    \label{fig:glon_d_pairlines}}
\end{figure*}
